---
title: "MRMR"
author: "Brian A. Fannin"
date: "14 July 2014"
output: beamer_presentation
fonttheme: professionalfonts
theme: Pittsburgh
colortheme: beaver
---

## What is MRMR?

- Flexible framework for reserving based on linear models
- Linear models = Chain ladder and Additive methods
- Emphasis on:
    - Ease of use
    - Multi-dimensional treatment of data
    - Visualization supports exploratory analysis and assessment of model fit
    
## Inspirations

- Leigh Halliwell
- Gelman & Hill, "Data Analysis Using Regression and Multilevel/Hierarchical Models"
- Jim Guszcza
- ggplot2, lubridate, sp

## Object structure

- OriginPeriod
- St*Measures
- Triangle
- TriangleModel
- TriangleProjection

## Common object properties

- May be thought of as data frames with constraints and metadata
- `[`, `$` accessors
- rbind, c, length
- New generics: 
    - write.excel
    - Grow

## OriginPeriod

```{r echo=FALSE, results='hide', message=FALSE}
library(knitr)
library(MRMR)
load("../MRMR/data/NAIC.rda")
load("../MRMR/data/smMulti.rda")
load("../MRMR/data/scmMulti.rda")
load("../MRMR/data/triMulti.rda")
```
```{r}
op = OriginPeriod(StartDate = as.Date("2001-01-01")
                  , NumPeriods = 10
                  , Period=as.period(1, "years")
                  , Type="Accident Year")
op$Moniker = paste0("AY ", year(op$StartDate))
```

## OriginPeriod (cont'd)

```{r echo=TRUE, eval=FALSE}
head(as.data.frame(op)[, 1:3])
```
```{r echo=FALSE, results='asis', size='huge', message=FALSE, tidy=FALSE, highlight=FALSE}
library(pander)
panderOptions("table.style", "rmarkdown")
df = as.data.frame(op)
dfDisplay = head(df)[, 1:3]
#dfDisplay$Period = as.character(dfDisplay$Period)
pander(head(dfDisplay))
```

## Non-standard OriginPeriods

```{r echo=TRUE, results='hide'}
startDates = seq(as.Date("2001/01/01")
                 , as.Date("2005/12/31"), by="6 months")
endDates = startDates + as.period(6, "months") - days(1)
op = OriginPeriod(startDates, endDates)
```

## Member access
```{r echo=TRUE, results='hide'}
op = OriginPeriod(seq(2001:2010))
op$Moniker = paste0("AY ", year(op$StartDate))
x = op[1]
y = op[2:3]
z = c(x, y)
y = op[c("AY 2004", "AY 2008")]
y = op$StartDate
y = op$Type
y = op$Moniker[3]
```

## St*Measures

- St*ticMeasures contain a set of observations associated with an OriginPeriod
- StaticMeasure objects do not vary over time
    + Earned Premium, Payroll, Number of autos
- StochasticMeasure objects vary over time
    + Paid loss, number of open claims

## St*Measures (cont'd)

- Contain a list of Levels of arbitray complexity
    + Each Level has a name and a set of attributes
    + Ex: Line level may have attributes "GL", "Motor", etc.
- Contain a set of Measures
    + StochasticMeasure will automatically compute incremental and prior measures

## Examples

All examples use NAIC data compiled by Glen Meyers and Peng Shi of ISO. Available at http://www.casact.org/research/index.cfm?fa=loss_reserves_data

## StaticMeasure construction

```{r eval=FALSE, results='hide'}
smWC = 
  StaticMeasure(OriginPeriod = op
                , Level=list(Company=companies
                             , Line="WC")
                , Measure=c("DirectEP", "NetEP")
                , Data=dfSubWC[dfSubWC$Lag == 1, ])
```

## StaticMeasure plot

```{r fig.height=6}
plot(smWC)
```

## StaticMeasure facet plot

```{r fig.height=6}
plot(smWC
     , FacetFormula=as.formula(Line+Company~Measure))
```

## StaticMeasure multidimensional plot

```{r fig.height=6}
plot(smMulti
     , FacetFormula=as.formula(Company~Line))
```

## StochasticMeasure - 1

```{r fig.height=6}
plot(UpperTriangle(scmWC)
     , Measure="CumulativeIncurred")
```

## StochasticMeasure - 2

```{r fig.height=6}
plot(UpperTriangle(scmWC)
     , Measure="CumulativeIncurred"
     , TimeAxis="EvaluationDate")
```

## Triangle

Very simple object. Composed of a StaticMeasure and a StochasticMeasure

```{r eval=FALSE, fig.height=6}
triWC = Triangle(smWC, scmWC, "Workers Comp Triangle")


plot(UpperTriangle(triWC)
     , Response="IncrementalPaid"
     , Predictor="PriorCumulativePaid")
```

## Triangle plot output

```{r echo=FALSE, fig.height=6}
plot(UpperTriangle(triWC), Predictor="PriorCumulativePaid", Response="IncrementalPaid")
```

## Triangle plot switch variables

```{r eval=FALSE}
plot(UpperTriangle(triWC)
     , Response="IncrementalPaid"
     , Predictor="PriorCumulativePaid")
```

```{r eval=FALSE, fig.height=6}
plot(UpperTriangle(triWC)
     , Response="IncrementalPaid"
     , Predictor="NetEP")
```

## Triangle plot - 2

```{r echo=FALSE, fig.height=6}
plot(UpperTriangle(triWC), Predictor="NetEP", Response="IncrementalPaid")
```

## TriangleModel

3 Model Types

- individual - Ignore the other data
- pooled - All data is equal
- blended - Weighted average of first two

## TriangleModel

```{r eval=FALSE}
PaidCLPooled = 
  TriangleModel(UpperTriangle(triWC)
                , Response = "IncrementalPaid"
                , Predictor = "PriorCumulativePaid"
                , ModelType = "pooled")
```

## Residual plots
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
PaidCLPooled = TriangleModel(UpperTriangle(triWC)
                             , Response = "IncrementalPaid"
                             , Predictor = "PriorCumulativePaid"
                             , ModelType = "pooled")

PaidCLIndividual = TriangleModel(UpperTriangle(triWC)
                             , Response = "IncrementalPaid"
                             , Predictor = "PriorCumulativePaid"
                             , ModelType = "individual")

PaidCLBlended = TriangleModel(UpperTriangle(triWC)
                              , Response = "IncrementalPaid"
                              , Predictor = "PriorCumulativePaid"
                              , Group = "Company"
                              , ModelType = "blended")
```
```{r warning=FALSE, fig.height=6}
PlotResiduals(PaidCLPooled)
```

## Residual plots 2

```{r warning=FALSE, fig.height=6}
PlotResiduals(PaidCLIndividual)
```

## Residual plots 3

```{r warning=FALSE, fig.height=6}
PlotResiduals(PaidCLBlended)
```

## Model comparison

```{r}
RMSE(PaidCLPooled)
RMSE(PaidCLIndividual)
RMSE(PaidCLBlended)
```


## Projected results

```{r}
projPaidCLPooled = 
  TriangleProjection(PaidCLPooled
                     , AsOfDate=as.Date("2006-12-31")
                     , MaxLag=10)
```

## Projected results

```{r echo=FALSE, results='asis'}
projPaidCLIndividual = TriangleProjection(PaidCLIndividual
                                      , AsOfDate=as.Date("2006-12-31")
                                      , MaxLag=10)

projPaidCLBlended = TriangleProjection(PaidCLBlended
                                      , AsOfDate=as.Date("2006-12-31")
                                      , MaxLag=10)

df.proj1 = as.data.frame(projPaidCLPooled@Projection)
df.proj2 = as.data.frame(projPaidCLIndividual@Projection)
df.proj3 = as.data.frame(projPaidCLBlended@Projection)

df.actual = as.data.frame(triWC)
companies = unique(df.actual$Company)
df.actual = subset(df.actual, Lag == 10)

df.actual = df.actual[, c("EvaluationDate", "IncrementalPaid")]
row.names(df.actual) = NULL

df.proj1 = df.proj1[, "IncrementalPaid"]
df.proj2 = df.proj2[, "IncrementalPaid"]
df.proj3 = df.proj3[, "IncrementalPaid"]

dfCompare = cbind(df.actual, df.proj1, df.proj2, df.proj3)
names(dfCompare)[2:5] = c("Actual", "Pooled", "Individual", "Blended")
dfCompare = subset(dfCompare, EvaluationDate > as.Date("1997-12-31"))

# panderOptions("table.style", "rmarkdown")
# pander(dfCompare[1:10, ])
```

```{r}
sqrt(sum((dfCompare$Actual - dfCompare$Pooled)^2))
sqrt(sum((dfCompare$Actual - dfCompare$Individual)^2))
sqrt(sum((dfCompare$Actual - dfCompare$Blended)^2))
```

## Model comparison
```{r echo=FALSE}
library(reshape2)
mdf = melt(dfCompare, id.vars=c("EvaluationDate", "Actual"), value.name="Predicted")
library(ggplot2)
plt = ggplot(mdf, aes(x=Actual, y=Predicted, group=variable, color=variable)) + geom_point() + geom_abline(intercept=0, slope=1)
plt
```

## R Stuff

- Lots of fun with S4 classes
- melt + ggplot = fun
- Multi-level models -> spoiled for choice
- XLConnect makes everyone happy

## What's next?

- Project on GitHub: https://github.com/PirateGrunt/MRMR
- Finalize for CRAN
    - Complete accessor and assignment functions
    - New generics methods for all objects
    - Vectorization of TriangleModel
    - Various of convenience functions
    - More visualizations!
    - Vignette
- Shiny application "shout"
- Full Bayesian model

## Who was that guy?

I work for The Redwoods Group. 

We're a niche provider of commercial insurance and we're dedicated to making the world a better place. 

We're small, but want to make a big impact. R is an ideal tool for us.
